---
title: "Project"
output: html_document
---

The goal of this project is to predict the manner in which they did the exercise. 
This is the "classe" variable in the training set. 

We should create a report describing how you built our model, how you used cross validation, what we think the expected out of sample error is, and why we made the choices we did. We will also use ourprediction model to predict 20 different test cases.

## 1.Preparing Data

The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

## 2.Loading Data

```{r}
training.csv<-read.csv("pml-training.csv", header = TRUE, na.strings=c("#DIV/0!","NA", ""))
testing.csv<-read.csv("pml-testing.csv", header = TRUE, na.strings=c("#DIV/0!","NA", ""))
dim(training.csv)
dim(testing.csv)
```

## 3.Cleaning Data

```{r}
# remove near zero variance
library(caret)
nearZero<-nearZeroVar(training.csv, saveMetrics=TRUE)
training.csv<-training.csv[, nearZero$nzv==FALSE]
# remove NAs
training.csv<-training.csv[ , colSums(is.na(training.csv))==0]
# remove not relevant columns
training.csv<-training.csv[,7:ncol(training.csv)]
dim(training.csv)
```

## 4.Random Forest

We create a model by using Random Forests.

```{r}
# split data
set.seed(33833)
inTrain<-createDataPartition(training.csv$classe,p=0.7,list=FALSE)
training<-training.csv[inTrain,]
testing<-training.csv[-inTrain,]

# train model
ctrl<-trainControl(method="cv", number=4, allowParallel=TRUE)
modFit<-train(classe ~ ., data=training, method="rf", trControl=ctrl)
modFit$finalModel

# in sample error
inError<-sum(diag(modFit$finalModel$confusion))/sum(modFit$finalModel$confusion)
inError

# cross validation
testingPredict<-predict(modFit, testing)
cm<-confusionMatrix(testingPredict,testing$classe)
cm

# out of sample error
outError<-sum(diag(cm$table))/sum(cm$table)
outError
```

The out of sample error is less than 1%.

## 5.Predicting 20 Test Cases

```{r}
predict20<-predict(modFit, testing.csv)
predict20

# create one file for each submission
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(predict20)
```


Thank you.
